Revisiting Short-Chain Fatty Acids and the Microbiota in Colorectal Cancer

### 2018-06-04 submission/manuscript.Rmd
smooth out  abstract

**Background.** Certain bacterial community members have been shown to increase colon tumorigenesis while other taxa have been associated with normal colons. The some of the taxa associated with normal colons can use fiber to produce short-chain fatty acids which have been shown to inhibit tumor growth in model systems. However, the data supporting the importance of SCFAs in humans is less certain. Here we test the hypothesis that short-chain fatty acid concentration changes in individuals with colorectal tumors.

@@

**Results.** No difference in short-chain fatty acid concentration was found between normal individuals and patients with adenoma or carcinoma (P-value `r ifelse(min(pull(hplc_groups_scfa_kruskal, bh)) > 0.15, invisible(">"), invisible("<"))` 0.15). There was no difference in the ability of classification models with or without short-chain fatty acids added to them in predicting adenoma or carcinoma (P-value `r ifelse(min(pull(disease_rf, bh)) > 0.05, invisible(">"), invisible("<"))` 0.05). Using metagenomic sequencing, there was also no difference in genes involved with short-chain fatty acid synthesis (P-value `r ifelse(min(pull(opf_select, bh)) > 0.70, invisible(">"), invisible("<"))` 0.70).

**Conclusions.** Although our data does not support the hypothesis that short-chain fatty acids decrease in individuals that have colorectal tumors, there may be specific scenarios where short-chain fatty acids may still be beneficial for treatment of colorectal cancer. Alternatively, there may be other metabolites that have not been thoroughly investigated that are more important to human colorectal cancer.

>>>>

**Background.** Colorectal cancer (CRC) is a growing health concern with a large percentage of the risk of developing disease being due to environmental factors. Certain bacterial community members have been associated to increase CRC while other taxa have been associated with normal colons. Some of the taxa associated with normal colons can use fiber to produce short-chain fatty acids (SCFAs) that can inhibit tumor growth in model systems. However, the data supporting the importance of SCFAs in human CRC is less certain. Here, we test the hypothesis that SCFA concentration changes in individuals with colorectal tumors.

@@

**Results.** No difference in SCFA concentration was found between normal individuals and patients with adenoma or carcinoma (P-value `r ifelse(min(pull(hplc_groups_scfa_kruskal, bh)) > 0.15, invisible(">"), invisible("<"))` 0.15). There was no difference in classification models with or without SCFAs in their ability to predict adenoma or carcinoma (P-value `r ifelse(min(pull(disease_rf, bh)) > 0.05, invisible(">"), invisible("<"))` 0.05). Using metagenomic sequencing, there was also no difference in genes involved with SCFA synthesis (P-value `r ifelse(min(pull(opf_select, bh)) > 0.70, invisible(">"), invisible("<"))` 0.70).

**Conclusions.** Although our data does not support the hypothesis that SCFAs decrease in individuals that have colorectal tumors, there may be specific scenarios where SCFAs may still be beneficial for treatment of CRC. Alternatively, there may be other metabolites that have not been thoroughly investigated that are more important to the development of human CRC.

### 2018-06-04 submission/manuscript.tex
smooth out  abstract

>>>>

\textbf{Background.} Colorectal cancer (CRC) is a growing health concern
with a large percentage of the risk of developing disease being due to
environmental factors. Certain bacterial community members have been
associated to increase CRC while other taxa have been associated with
normal colons. Some of the taxa associated with normal colons can use
fiber to produce short-chain fatty acids (SCFAs) that can inhibit tumor
growth in model systems. However, the data supporting the importance of
SCFAs in human CRC is less certain. Here, we test the hypothesis that
SCFA concentration changes in individuals with colorectal tumors.

\textbf{Methods.} We analyzed a cross-sectional (n=490) and pre- and
post-treatment (n=67) group for their concentrations of acetate,
butyrate, and propionate. Analysis also included tumor classification
models using Random Forest, imputed gene relative abundance with
PICRUSt, and metagenomic sequencing on a subset (n=85) of the total
cross-sectional group.

\textbf{Results.} No difference in SCFA concentration was found between
normal individuals and patients with adenoma or carcinoma (P-value
\textgreater{} 0.15). There was no difference in classification models
with or without SCFAs in their ability to predict adenoma or carcinoma
(P-value \textgreater{} 0.05). Using metagenomic sequencing, there was
also no difference in genes involved with SCFA synthesis (P-value
\textgreater{} 0.70).

\textbf{Conclusions.} Although our data does not support the hypothesis
that SCFAs decrease in individuals that have colorectal tumors, there
may be specific scenarios where SCFAs may still be beneficial for
treatment of CRC. Alternatively, there may be other metabolites that
have not been thoroughly investigated that are more important to the
development of human CRC.
